The role of advection versus diffusion in intraventricular transport

### 2025-02-03 paper/main.tex
Updates from discussion just now.

% Title
\title{The role of cilia-mediated cerebrospinal fluid flow in intraventricular transport in zebrafish embryos}
\author[1,*]{Halvor Herlyng}
\author[1]{Ada J.~Ellingsrud}
\author[1]{Miroslav Kuchta} 
\author[X]{Inyoung Jeong} % OK?
\author[1,3,+]{Marie E.~Rognes} % OK order?
\author[2,+]{Nathalie Jurisch-Yaksi}
\affil[1]{Department for Numerical Analysis and Scientific Computing, Simula Research Laboratory, Oslo, Norway}
\affil[2]{Department of Clinical and Molecular Medicine, Norwegian University of Science and Technology, Trondheim, Norway}
\affil[3]{K.~G.~Jebsen Center for Brain Fluid Research}
\affil[*]{\href{mailto:hherlyng@simula.no}{hherlyng@simula.no}}
\affil[+]{these authors contributed equally to this work}
% Optional PDF information
\ifpdf
\hypersetup{
  pdftitle={zebrafish-cilia-csf},
  pdfauthor={H.~Herlyng \textit{et al.}}
}
\fi

@@

\subsection*{Image-based computational geometries of zebrafish brain ventricles}
We generated a 3D representation of the brain ventricular walls using confocal imaging data of a zebrafish embryo injected intraventricularly with a 70 kDa dye at 2 dpf~\cite{Olstad2019CiliaryDevelopment}. From the imaging data, we constructed a computational mesh of the ventricles (\cref{fig:figure1}, panel \textbf{d}). The interior volume was meshed with fTetWild~\cite{Hu2020FastWild}. The \emph{standard} mesh consisted of 132 134 tetrahedral cells and had a maximal (minimal) \mer{edge length} of 14.9 $\mu$m (4.68 $\mu$m). At 2 dpf, the ventricular system consists mainly of 3 cavities: the telencephalic  (anterior), the di-/mesencephalic  (middle) and the rhombencephalic (posterior) ventricles, which are connected by ducts. Parts of the ventricular surfaces were marked to be lined with motile cilia~\cite{Olstad2019CiliaryDevelopment} (\cref{fig:figure1}\textbf{d}, pink markers).
\mer{MER: @HH: Check if you mean edge length or mesh cell size here.}
\begin{figure}%[hbt!]
    \centering
    \includegraphics[width=\linewidth]{graphics/figure1.png}
    \caption{\lyng{Experimenting with velocity colorbars, thoughts on the two options? Or have two separate as it is now?}}
    \label{fig:figure1}
\end{figure}

@@

The beating motion of the cilia generates CSF flow with steady rotational structures~\cite{Olstad2019CiliaryDevelopment} within the brain ventricles (\cref{fig:figure1}, panels \textbf{a}--\textbf{c}). In addition, cardiac pulsations induces pulsatile CSF flow. We model this flow of CSF in the brain ventricles by the time-dependent, incompressible Stokes equations which read as follows. Find the CSF velocity $\uu = \uu(\xx, t) = (u_x(\xx, t), u_y(\xx, t), u_z(\xx, t))$ and the CSF pressure $p = p(\xx, t)$ for $\xx \in \Omega$ and time $t>0$, such that
\begin{subequations}
    \begin{align}
      \rho \pdifft{\uu} - \nabla \cdot \bsig &= \mathbf{0}
      \quad \text{in } \Omega,
      \label{eq:stokes_eq_mom}\\
      \nabla \cdot \uu &= 0
      \quad  \text{in } \Omega,
  \end{align}
  \label{eq:stokes_eqs}% 
\end{subequations}%
where the stress tensor is defined as $\bsig(\uu, p) = 2\mu\beps(\uu) - p\mathbf{I}$ and $\beps(\uu) = \frac{1}{2}\left(\nabla \uu + (\nabla\uu)^T\right)$ is the strain-rate tensor. Bold-face characters denote vectors or tensors, and $\mathbf{I}$ is the identity tensor in three dimensions. Furthermore, $\rho$ is the CSF density and $\mu$ is the dynamic viscosity of the CSF. We introduce the tangential traction $\bsigpar$ as the tangential component of the traction $\hat{\bsig}=\bsig\nn$:
\begin{equation*}
    \bsigpar = P_{\nn}(\hat{\bsig}) = (\mathbf{I} - \nn\otimes\nn)\hat{\bsig},
\end{equation*}
where $\nn$ is the outer unit normal of the boundary surface $\Gamma = \partial\Omega$, and $P_{\nn}(\rr)$ is the tangential  projection of a vector $\rr$ onto $\Gamma$. We consider three types of boundary conditions:  (i) \emph{tangential traction} (no flow normal to the boundary $\uu \cdot \nn = 0$ with a force $\btau$ applied tangentially on the boundary $\bsigpar = \btau$), (ii) \emph{normal pressure} (a pressure $\Tilde{p}$ applied in the normal direction: $(\mu\nabla\uu - p\mathbf{I})\nn = \Tilde{p}(t)\nn$), and (iii) \emph{free slip} (no flow normal to the boundary $\uu \cdot \nn = 0$ and no tangential forces $\bsigpar = {\bf 0}$). Unless noted otherwise, the free-slip condition is applied on the boundary. We use the initial condition\mer{s} $\uu(\xx, 0)=\mathbf{0}$ \mer{and $p(\xx, t) = 0$} for all $\xx\in\Omega$. \mer{MER: I don't think you use an initial condition on the pressure. Please check and remove if so. You can also just write: ``We assume that the system starts at rest ($\uu(t = 0) = \mathbf{0}$)''.}

@@

\subsection*{Computational model of solute transport after photoconversion}
We simulate transport of photoconverted proteins within the brain ventricles by modeling transport of a solute with concentration $c(\xx, t)$ for $\xx\in\Omega$ and time $t > 0$ by the advection-diffusion equation
\begin{equation}
    \pdifft{c} + \nabla\cdot\JJ = 0 \quad \text{in } \Omega,
    \label{eq:adv_diff_strong}
\end{equation}
where the total concentration flux $\mathbf{J}$ is assumed to consist of an advective and a diffusive flux:
\begin{equation*}
    \JJ = c\uu - D\nabla c .
\end{equation*}
Here, $\uu$ is the CSF velocity field, governed by the Stokes equations~\eqref{eq:stokes_eqs}, and $D$ is the molecular diffusion coefficient of the solute. Initially, we set $c(\xx, 0)=0$ for all $\xx\in\Omega$. The ventricular walls are assumed to be impermeable (admitting \emph{no flux} i.e.~$\JJ\cdot\nn=0$), except at the anterior/posterior boundary $\Gamma_p$. \fixme{There, we consider either prescribed \emph{influx} in the form of a time-periodic condition on the total flux $(\JJ\cdot\nn)|_{\mathrm{in}}^{n+1}=(\JJ\cdot\nn)|_{\mathrm{out}}^{n}$ and/or \emph{outflux} with zero diffusive flux $D\nabla c\cdot\nn=0$.}
\mer{MER: Could we make this section about the boundary conditions easier to read? I've had a go at it, please check.}

\fixme{Additionally, when considering pulsatile motion, the in- and outflow on the pressure boundary $\Gp$ must be accounted for. We split $\Gp$ in two parts; $\Gin$ denotes the part of $\Gp$ where $\uu\cdot\nn < 0$; $\Gout$ denotes the part where $\uu\cdot\nn > 0$. On $\Gin$, a total flux boundary condition was set: the inflow flux at a given timestep was set equal to the outflow boundary flux at the previous timestep. This was done to approximate a periodic boundary condition. On $\Gout$, we considered the diffusive flux normal to the boundary to be zero.}
\mer{MER: I moved this paragraph from the section below, because it seemed to belong here. But needs edits.}

@@

Motivated by an interest in molecules that are significant in neural development, we study the transport of molecules resembling Dendra2 Fluorescent Protein (Dendra2), Starmaker+Green Fluorescent Protein (STM+GFP) and Exosomes. To estimate the diffusion coefficients of these, when and if not available in the literature, we use the Stokes--Einstein relation:
\begin{equation}
    D = \frac{k_B T}{6\pi \mu R} ,
    \label{eq:D_stokes_einstein}
\end{equation}
which is an expression for the diffusion coefficient of a spherical molecule suspended in a viscous fluid~\cite{Einstein1905UberTeilchen}. In~\eqref{eq:D_stokes_einstein}, $k_B = \num{1.38e-23}$ J/K is the Boltzmann constant, $T = 310$ K is the absolute temperature, $\mu$ is the dynamic viscosity of the solvent (in this case the CSF), and $R$ is the radius of the molecule. For the exosomes, we calculate $D$ using~\eqref{eq:D_stokes_einstein} with $R=200$ nm. This radius covers the slowest diffusing exosomes, since the smaller exosomes would have diffusion coefficients similar to the other molecules. For STM+GFP, we estimated the diffusion coefficient by extrapolating the experimentally observed diffusion coefficient of GFP~\cite{Swaminathan1997PhotobleachingDiffusion, Potma2001ReducedCells}, assuming that if $D$ scales with $1/R$ according to~\eqref{eq:D_stokes_einstein}, $D$ would scale with the cube root of the molecule mass~\cite{Goodhill1997DiffusionGuidance}. The resulting diffusion coefficients used in the simulations are summarized in \cref{tab:parameter_table}.

@@

%% The molecular diffusion coefficients were estimated in part based on values reported in the literature, and in part via extrapolation with the Stokes--Einstein relation (\cref{tab:diff_coeff_table}). The Stokes-Einstein relation
\begin{table}[h!]
    \centering
    \label{tab:diff_coeff_table}
    \begin{tabular}{l|ccc}
        \toprule
        Molecule & Mass [kDa] & $D \ [\mathrm{m^2/s}]$ & Reference \\ 
        \midrule 
        Dendra2 Fluorescent Protein & 25.6~\cite{Gurskaya2006EngineeringLight} & $\num{1.15e-10}$  & \cite{GuraSadovsky2017MeasurementExpansion}\\
        GFP (saline aqueous solution) & 26.9~\cite{UniProt2024GFPUniProtKB} & $\num{8.70e-11}$ & \cite{Swaminathan1997PhotobleachingDiffusion, Potma2001ReducedCells} \\
        STM+GFP & 66.2~\cite{UniProt2024StmUniProtKB} + 26.9~\cite{UniProt2024GFPUniProtKB} & $\num{5.75e-11}$ & * \\
        Exosomes (radius of 200 nm) & -- & $\num{1.63e-12}$ & \textdagger \\
        \bottomrule
    \end{tabular}
    \caption{Molecular mass and diffusion coefficients of simulated solutes. STM and GFP denote Starmaker Protein and Green Fluorescent Protein, respectively. (*) The value was extrapolated based on the value of $D$ for GFP reported in the literature~\cite{Swaminathan1997PhotobleachingDiffusion, Potma2001ReducedCells}, assuming that $D$ scales with the cube root of the molecule mass. (\textdagger) The value was estimated using the Stokes-Einstein relation~\eqref{eq:D_stokes_einstein}.}
\end{table}

@@

\subsection*{Physical parameters}
%\label{subsec:parameters}
The CSF density $\rho$ and dynamic viscosity $\mu$ were set following Bloomfield \textit{et al.}~\cite{Bloomfield1998EffectsFluid}~(\cref{tab:parameter_table}). We considered large, medium and small-sized molecules by three different diffusion coefficients $D_1, D_2$ and $D_3$ (\Cref{tab:parameter_table}). Experimental data~\cite{Olstad2019CiliaryDevelopment} indicate \fixme{velocities} proximate to the ventricular walls of $27.4 \pm 5.4$ \textmu m/s and $2.6 \pm 0.6$ \textmu m/s in the dorsal and ventral regions of the diencephalic ventricle, respectively. In the telencephalic ventricle, the velocities were estimated at $4.7 \pm 1.3$ \textmu m/s. We calibrated $\tau$ and $A$ heuristically by comparing these experimental velocity values with the simulated, maximum velocity magnitude. In addition, particle tracking data of pulsatile motion~(\cref{sec:appC_particle_transport}) was used in the calibration process to determine the relative contributions of the cardiac pulsations, as compared to the cilia-driven flow, to the velocity magnitude.
\mer{MER: What velocities? Net velocities? Velocity magnitude aka speed? Also, haven't you said this part about the experimental data already? Either place the details there, or all here. I suggest there.}
\mer{MER: I suggest you move the parts relevant for cilia/cardiac flow to where you mentioned it first, and then merge this section with the Estimation of diffusion .... Just refer to \cref{tab:parameter_table}. Actually, unsure if you need it. You say everything in the text or in the other Table right?}
\begin{table}[h!]
    \centering
    \begin{tabular}{lcccc}
    \toprule
        Parameter & Symbol & Value & Unit & Reference \\
        \midrule
        Fluid density & $\rho$ & $1000$ & $\mathrm{kg/m^3}$ & \cite{Bloomfield1998EffectsFluid} \\
        Dynamic viscosity & $\mu$ & $\num{6.97e-04}$ & $\mathrm{Pa\cdot s}$ & \cite{Bloomfield1998EffectsFluid} \\
        Cardiac frequency & $f$ & 2.22 & Hz &  \cite{Olstad2019CiliaryDevelopment}\\
        Angular cardiac frequency & $\omega = 2\pi f$ & 6.97 & rad/s & \cite{Olstad2019CiliaryDevelopment}\\
        Cilia tangential traction & $\tau$ & \num{6.5e-04} & Pa & * \\
        Pressure amplitude & $A$ & \num{1.5e-03} & Pa & * \\
        %Ref. frame motion amplitude & $A_{\mathrm{ref}}$ & \num{9.19e-04} & m/s & *\\
        Exosome diffusion coefficient & $D_1$ & $\num{1.63e-12}$ & $\mathrm{m^2/s}$ & \textdagger\\
        STM+GFP diffusion coefficient  & $D_2$ & $\num{5.75e-11}$ & $\mathrm{m^2/s}$ & \textdagger\\
        Dendra2 diffusion coefficient  & $D_3$ & $\num{1.15e-10}$ & $\mathrm{m^2/s}$ & \textdagger\\
    \bottomrule
    \end{tabular}
    \caption{Physical parameters of the flow and transport models. (*) Calibrated based on experimental data~\cite{Olstad2019CiliaryDevelopment}. (\textdagger) Estimated based on experimental data~\cite{GuraSadovsky2017MeasurementExpansion, Swaminathan1997PhotobleachingDiffusion, Potma2001ReducedCells}.}
    \label{tab:parameter_table}
\end{table}

\subsection*{Computational modelling scenarios}
%\label{subsec:model_versions}
To study the impact of the motile cilia and the cardiac pulsations on the CSF flow, we consider the following computational flow scenarios, each predicting a CSF flow velocity $\uu$ and CSF pressure $p$.
\begin{itemize}
    \item A \emph{baseline} flow model representing a wild-type zebrafish including contributions both from the cilia forces and the cardiac pulsations. In this scenario, we apply the tangential traction $\btau$ on the cilia boundary $\Gc$ and prescribe the normal pressure $\Tilde{p}$ on the anterior/posterior boundary $\Gp$. % Model 0

@@

\mer{MER: Let's discuss the paragraphs below and see where they fit best. I think we can drop the first paragraph (you said this already). I find the second paragraph to be somewhat confusing (low signal-to-noise ratio).}
When simulating transport of solute in the ventricles, we consider the same six different regions of interest that were considered in the photoconversion experiments. In region number one (the photoconversion region), we use the function~\eqref{eq:injection_curve} to enforce the concentration such that it represents the experimental data of photoconversion of proteins. We consider the three different diffusion coefficients introduced in previous sections to simulate and compare results for transport of molecules of varying sizes.

@@

\mer{One sentence on why? - what do we not know - what do we try to
  find out?} To characterize CSF flow patterns predicted by
computational fluid dynamics within the brain ventricles, we begin by
considering the simulated CSF velocity and pressure when driven by
both cilia motion and cardiac pressure pulsations. These forces induce
\fixme{persistent} flow features, with vortex-structures in the
anterior and middle ventricles and \fixme{unidirectional} flow in the
posterior ventricle (\Cref{fig:figure1}\textbf{e},
\fixme{Supplementary video 1}). These characteristic patterns suggest
that the flow in the anterior and middle ventricles are dominated by
the cilia contribution, while the cardiac pressure pulsations dominate
within the posterior ventricle. Over one cardiac cycle, the CSF flow
speed reaches up to 27.7 \textmu m/s, dorsally in the posterior part
of the middle ventricle. This peak is attained at the time instant
coinciding with the time of peak caudal cardiac pulsatile flow, as a
result of both the traction induced by the cilia and the pulsatile
pressure-driven flow being aligned with the rostrocaudal ($x-$)
axis. Using this maximum velocity magnitude and choosing the height of
the middle ventricle ($110$ \textmu m) as a characteristic length
scale, yields a Reynolds number $\mathrm{Re} \approx 0.004$,
justifying the assumption of Stokes flow. To further examine the
effect of the cilia, we next consider a no-cilia model variation. In
this scenario, the simulated CSF flow is pulsatile and
\fixme{unidirectional} (\cref{fig:figure1}g), with a peak velocity
magnitude of \fixme{$4.8$ \textmu m/s}. On the other hand, the
vortices (re)appear when considering a variation with cilia but no
cardiac contribution (\cref{fig:figure1}h), in which case the highest
velocities ($26.5$ \textmu m/s) also occur in the vicinity of the
cilia. Thus, the bidirectional traction induced by motile cilia
inherently leads to a partial compartmentalization of the ventricular
system.

>>>>

% Title
\title{The role of advection versus diffusion in intraventricular transport}
\author[1,*]{Halvor Herlyng}
\author[1]{Ada J.~Ellingsrud}
\author[1]{Miroslav Kuchta} 
\author[X]{Inyoung Jeong} % OK?
\author[1,3,+]{Marie E.~Rognes} % OK order?
\author[2,+]{Nathalie Jurisch-Yaksi}
\affil[1]{Department for Numerical Analysis and Scientific Computing, Simula Research Laboratory, Oslo, Norway}
\affil[2]{Department of Clinical and Molecular Medicine, Norwegian University of Science and Technology, Trondheim, Norway}
\affil[3]{K.~G.~Jebsen Center for Brain Fluid Research}
\affil[*]{\href{mailto:hherlyng@simula.no}{hherlyng@simula.no}}
\affil[+]{these authors contributed equally to this work}
% Optional PDF information
\ifpdf
\hypersetup{
  pdftitle={zebrafish-cilia-csf},
  pdfauthor={H.~Herlyng \textit{et al.}}
}
\fi

@@

\subsection*{Image-based computational geometries of zebrafish brain ventricles}
We generated a 3D representation of the brain ventricular walls using confocal imaging data of a zebrafish embryo injected intraventricularly with a 70 kDa dye at 2 dpf~\cite{Olstad2019CiliaryDevelopment}. From the imaging data, we constructed a computational mesh of the ventricles (\cref{fig:figure1}, panel \textbf{d}). The interior volume was meshed with fTetWild~\cite{Hu2020FastWild}. The \emph{standard} mesh consisted of 132 134 tetrahedral cells and had a maximal (minimal) edge length of 14.9 $\mu$m (4.68 $\mu$m). At 2 dpf, the ventricular system consists mainly of 3 cavities: the telencephalic  (anterior), the di-/mesencephalic  (middle) and the rhombencephalic (posterior) ventricles, which are connected by ducts. Parts of the ventricular surfaces were marked to be lined with motile cilia~\cite{Olstad2019CiliaryDevelopment} (\cref{fig:figure1}d, pink markers).
\begin{figure}%[hbt!]
    \centering
    \includegraphics[width=\linewidth]{graphics/figure1.png}
    \caption{\lyng{Experimenting with velocity colorbars, thoughts on the two options? Or have two separate as it is now?}}
    \label{fig:figure1}
\end{figure}

@@

The beating motion of the cilia generates CSF flow with steady rotational structures~\cite{Olstad2019CiliaryDevelopment} within the brain ventricles (\cref{fig:figure1}, panels \textbf{a}--\textbf{c}). In addition, cardiac pulsations induces pulsatile CSF flow. We model this flow of CSF in the brain ventricles by the time-dependent, incompressible Stokes equations which read as follows. Find the CSF velocity $\uu = \uu(\xx, t) = (u_x(\xx, t), u_y(\xx, t), u_z(\xx, t))$ and the CSF pressure $p = p(\xx, t)$ for $\xx \in \Omega$ and time $t>0$, such that
\begin{subequations}
    \begin{align}
      \rho \pdifft{\uu} - \nabla \cdot \bsig &= \mathbf{0}
      \quad \text{in } \Omega,
      \label{eq:stokes_eq_mom}\\
      \nabla \cdot \uu &= 0
      \quad  \text{in } \Omega,
  \end{align}
  \label{eq:stokes_eqs}% 
\end{subequations}%
where the stress tensor is defined as $\bsig(\uu, p) = 2\mu\beps(\uu) - p\mathbf{I}$ and $\beps(\uu) = \frac{1}{2}\left(\nabla \uu + (\nabla\uu)^T\right)$ is the strain-rate tensor. Bold-face characters denote vectors or tensors, and $\mathbf{I}$ is the identity tensor in three dimensions. Furthermore, $\rho$ is the CSF density and $\mu$ is the dynamic viscosity of the CSF. We introduce the tangential traction $\bsigpar$ as the tangential component of the traction $\hat{\bsig}=\bsig\nn$:
\begin{equation*}
    \bsigpar = P_{\nn}(\hat{\bsig}) = (\mathbf{I} - \nn\otimes\nn)\hat{\bsig},
\end{equation*}
where $\nn$ is the outer unit normal of the boundary surface $\Gamma = \partial\Omega$, and $P_{\nn}(\rr)$ is the tangential  projection of a vector $\rr$ onto $\Gamma$. We consider three types of boundary conditions:  (i) \emph{tangential traction} (no flow normal to the boundary $\uu \cdot \nn = 0$ with a force $\btau$ applied tangentially on the boundary $\bsigpar = \btau$), (ii) \emph{normal pressure} (a pressure $\Tilde{p}$ applied in the normal direction: $(\mu\nabla\uu - p\mathbf{I})\nn = \Tilde{p}(t)\nn$), and (iii) \emph{free slip} (no flow normal to the boundary $\uu \cdot \nn = 0$ and no tangential forces $\bsigpar = {\bf 0}$). Unless noted otherwise, the free-slip condition is applied on the boundary. We assume that the system starts at rest ($\uu(t = 0) = \mathbf{0}$).

@@

\subsection*{Computational model of solute transport after photoconversion}
We simulate transport of photoconverted proteins within the brain ventricles by modeling transport of a solute with concentration $c(\xx, t)$ for $\xx\in\Omega$ and time $t > 0$ by the advection-diffusion equation
\begin{equation}
    \pdifft{c} + \nabla\cdot\JJ = 0 \quad \text{in } \Omega,
    \label{eq:adv_diff_strong}
\end{equation}
where the total concentration flux $\mathbf{J}$ is assumed to consist of an advective and a diffusive flux:
\begin{equation*}
    \JJ = c\uu - D\nabla c .
\end{equation*}
Here, $\uu$ is the CSF velocity field, governed by the Stokes equations~\eqref{eq:stokes_eqs}, and $D$ is the molecular diffusion coefficient of the solute. Initially, we set $c(\xx, 0)=0$ for all $\xx\in\Omega$. The ventricular walls are assumed to be impermeable (admitting \emph{no flux} i.e.~$\JJ\cdot\nn=0$), except at the anterior/posterior boundary $\Gamma_p$. \fixme{At the anterior/posterior boundary we allow the solute concentration to be advected into and out of the domain in a manner that ensures conservation of mass (see \Cref{sec:appendix:X} for details).}
\mer{MER: @Halvor: could you have a go at moving the precise description to the Appendix?}
\fixme{There, we consider either prescribed \emph{influx} in the form of a time-periodic condition on the total flux $(\JJ\cdot\nn)|_{\mathrm{in}}^{n+1}=(\JJ\cdot\nn)|_{\mathrm{out}}^{n}$ and/or \emph{outflux} with zero diffusive flux $D\nabla c\cdot\nn=0$.}
\fixme{Additionally, when considering pulsatile motion, the in- and outflow on the pressure boundary $\Gp$ must be accounted for. We split $\Gp$ in two parts; $\Gin$ denotes the part of $\Gp$ where $\uu\cdot\nn < 0$; $\Gout$ denotes the part where $\uu\cdot\nn > 0$. On $\Gin$, a total flux boundary condition was set: the inflow flux at a given timestep was set equal to the outflow boundary flux at the previous timestep. This was done to approximate a periodic boundary condition. On $\Gout$, we considered the diffusive flux normal to the boundary to be zero.}
%\mer{MER: I moved this paragraph from the section below, because it seemed to belong here. But needs edits.}

@@

Motivated by an interest in molecules that are significant in neural development, we study the transport of molecules resembling Dendra2 Fluorescent Protein (Dendra2), Starmaker+Green Fluorescent Protein (STM+GFP) and Exosomes. To estimate the diffusion coefficients of these, when and if not available in the literature, we use the Stokes--Einstein relation:
\begin{equation}
    D = \frac{k_B T}{6\pi \mu R} ,
    \label{eq:D_stokes_einstein}
\end{equation}
which is an expression for the diffusion coefficient of a spherical molecule suspended in a viscous fluid~\cite{Einstein1905UberTeilchen}. In~\eqref{eq:D_stokes_einstein}, $k_B = \num{1.38e-23}$ J/K is the Boltzmann constant, $T = 310$ K is the absolute temperature, $\mu$ is the dynamic viscosity of the solvent (in this case the CSF), and $R$ is the radius of the molecule. For the exosomes, we calculate $D$ using~\eqref{eq:D_stokes_einstein} with $R=200$ nm. This radius covers the slowest diffusing exosomes, since the smaller exosomes would have diffusion coefficients similar to the other molecules. For STM+GFP, we estimated the diffusion coefficient by extrapolating the experimentally observed diffusion coefficient of GFP~\cite{Swaminathan1997PhotobleachingDiffusion, Potma2001ReducedCells}, assuming that if $D$ scales with $1/R$ according to~\eqref{eq:D_stokes_einstein}, $D$ would scale with the cube root of the molecule mass~\cite{Goodhill1997DiffusionGuidance}. The resulting diffusion coefficients used in the simulations are summarized in \cref{tab:diff_coeff_table}.

@@

%% The molecular diffusion coefficients were estimated in part based on values reported in the literature, and in part via extrapolation with the Stokes--Einstein relation (\cref{tab:diff_coeff_table}). The Stokes-Einstein relation
\begin{table}[h!]
    \centering
    \begin{tabular}{l|ccc}
        \toprule
        Molecule & Mass [kDa] & $D \ [\mathrm{m^2/s}]$ & Reference \\ 
        \midrule 
        Dendra2 Fluorescent Protein & 25.6~\cite{Gurskaya2006EngineeringLight} & $\num{1.15e-10}$  & \cite{GuraSadovsky2017MeasurementExpansion}\\
        GFP (saline aqueous solution) & 26.9~\cite{UniProt2024GFPUniProtKB} & $\num{8.70e-11}$ & \cite{Swaminathan1997PhotobleachingDiffusion, Potma2001ReducedCells} \\
        STM+GFP & 66.2~\cite{UniProt2024StmUniProtKB} + 26.9~\cite{UniProt2024GFPUniProtKB} & $\num{5.75e-11}$ & * \\
        Exosomes (radius of 200 nm) & -- & $\num{1.63e-12}$ & \textdagger \\
        \bottomrule
    \end{tabular}
    \caption{Molecular mass and diffusion coefficients of simulated solutes. STM and GFP denote Starmaker Protein and Green Fluorescent Protein, respectively. (*) The value was extrapolated based on the value of $D$ for GFP reported in the literature~\cite{Swaminathan1997PhotobleachingDiffusion, Potma2001ReducedCells}, assuming that $D$ scales with the cube root of the molecule mass. (\textdagger) The value was estimated using the Stokes-Einstein relation~\eqref{eq:D_stokes_einstein}.}
    \label{tab:diff_coeff_table}
\end{table}

@@

\subsection*{Physical parameters}
%\label{subsec:parameters}
The CSF density $\rho$ and dynamic viscosity $\mu$ were set following Bloomfield \textit{et al.}~\cite{Bloomfield1998EffectsFluid}~(\cref{tab:parameter_table}). We considered large, medium and small-sized molecules by three different diffusion coefficients $D_1, D_2$ and $D_3$ (\Cref{tab:parameter_table}). Experimental data~\cite{Olstad2019CiliaryDevelopment} indicate \fixme{velocities} proximate to the ventricular walls of $27.4 \pm 5.4$ \textmu m/s and $2.6 \pm 0.6$ \textmu m/s in the dorsal and ventral regions of the diencephalic ventricle, respectively. In the telencephalic ventricle, the velocities were estimated at $4.7 \pm 1.3$ \textmu m/s. We calibrated $\tau$ and $A$ heuristically by comparing these experimental velocity values with the simulated, maximum velocity magnitude. In addition, particle tracking data of pulsatile motion~(\cref{sec:appC_particle_transport}) was used in the calibration process to determine the relative contributions of the cardiac pulsations, as compared to the cilia-driven flow, to the velocity magnitude.
\mer{MER: What velocities? Net velocities? Velocity magnitude aka speed? Also, haven't you said this part about the experimental data already? Either place the details there, or all here. I suggest there.}
\mer{MER: I suggest you introduce parameter values when introduced, and move the parts relevant for cilia/cardiac flow to where you mentioned it first, and then merge this section with the Estimation of diffusion .... Then remove this section and \Cref{tab:parameter_table}.}
\begin{table}[h!]
    \centering
    \begin{tabular}{lcccc}
    \toprule
        Parameter & Symbol & Value & Unit & Reference \\
        \midrule
        Fluid density & $\rho$ & $1000$ & $\mathrm{kg/m^3}$ & \cite{Bloomfield1998EffectsFluid} \\
        Dynamic viscosity & $\mu$ & $\num{6.97e-04}$ & $\mathrm{Pa\cdot s}$ & \cite{Bloomfield1998EffectsFluid} \\
        Cardiac frequency & $f$ & 2.22 & Hz &  \cite{Olstad2019CiliaryDevelopment}\\
        Angular cardiac frequency & $\omega = 2\pi f$ & 6.97 & rad/s & \cite{Olstad2019CiliaryDevelopment}\\
        Cilia tangential traction & $\tau$ & \num{6.5e-04} & Pa & * \\
        Pressure amplitude & $A$ & \num{1.5e-03} & Pa & * \\
        %Ref. frame motion amplitude & $A_{\mathrm{ref}}$ & \num{9.19e-04} & m/s & *\\
        Exosome diffusion coefficient & $D_1$ & $\num{1.63e-12}$ & $\mathrm{m^2/s}$ & \textdagger\\
        STM+GFP diffusion coefficient  & $D_2$ & $\num{5.75e-11}$ & $\mathrm{m^2/s}$ & \textdagger\\
        Dendra2 diffusion coefficient  & $D_3$ & $\num{1.15e-10}$ & $\mathrm{m^2/s}$ & \textdagger\\
    \bottomrule
    \end{tabular}
    \caption{Physical parameters of the flow and transport models. (*) Calibrated based on experimental data~\cite{Olstad2019CiliaryDevelopment}. (\textdagger) Estimated based on experimental data~\cite{GuraSadovsky2017MeasurementExpansion, Swaminathan1997PhotobleachingDiffusion, Potma2001ReducedCells}.}
    \label{tab:parameter_table}
\end{table}

\subsection*{Computational flow modelling scenarios}
%\label{subsec:model_versions}
To study the impact of the motile cilia and the cardiac pulsations on the CSF flow, we consider the following computational flow scenarios, each predicting a CSF flow velocity $\uu$ and CSF pressure $p$.
\begin{itemize}
    \item A \emph{baseline} flow model representing a wild-type zebrafish including contributions both from the cilia forces and the cardiac pulsations. In this scenario, we apply the tangential traction $\btau$ on the cilia boundary $\Gc$ and prescribe the normal pressure $\Tilde{p}$ on the anterior/posterior boundary $\Gp$. % Model 0

@@

\mer{MER: @Halvor I suggest you move these two paragraphs into their respective Results sections (Results 2 and 5). Putting the bullet points above at the end of the Stokes equations is a great idea!}
When simulating transport of solute in the ventricles, we consider the same six different regions of interest that were considered in the photoconversion experiments. In region number one (the photoconversion region), we use the function~\eqref{eq:injection_curve} to enforce the concentration such that it represents the experimental data of photoconversion of proteins. We consider the three different diffusion coefficients introduced in previous sections to simulate and compare results for transport of molecules of varying sizes.

@@

\mer{One sentence on why? - what do we not know - what do we try to
  find out?} \mer{@Halvor: If you want, try to revamp this paragraph
  as suggested by Nathalie. First cilia. Then pulsatile. Then
  combined.} To characterize CSF flow patterns predicted by
computational fluid dynamics within the brain ventricles, we begin by
considering the simulated CSF velocity and pressure when driven by
both cilia motion and cardiac pressure pulsations. These forces induce
\fixme{persistent} flow features, with vortex-structures in the
anterior and middle ventricles and \fixme{unidirectional} flow in the
posterior ventricle (\Cref{fig:figure1}\textbf{e},
\fixme{Supplementary video 1}). These characteristic patterns suggest
that the flow in the anterior and middle ventricles are dominated by
the cilia contribution, while the cardiac pressure pulsations dominate
within the posterior ventricle. Over one cardiac cycle, the CSF flow
speed reaches up to 27.7 \textmu m/s, dorsally in the posterior part
of the middle ventricle. This peak is attained at the time instant
coinciding with the time of peak caudal cardiac pulsatile flow, as a
result of both the traction induced by the cilia and the pulsatile
pressure-driven flow being aligned with the rostrocaudal ($x-$)
axis. Using this maximum velocity magnitude and choosing the height of
the middle ventricle ($110$ \textmu m) as a characteristic length
scale, yields a Reynolds number $\mathrm{Re} \approx 0.004$,
justifying the assumption of Stokes flow. To further examine the
effect of the cilia, we next consider a no-cilia model variation. In
this scenario, the simulated CSF flow is pulsatile and
\fixme{unidirectional} (\cref{fig:figure1}g), with a peak velocity
magnitude of \fixme{$4.8$ \textmu m/s}. On the other hand, the
vortices (re)appear when considering a variation with cilia but no
cardiac contribution (\cref{fig:figure1}h), in which case the highest
velocities ($26.5$ \textmu m/s) also occur in the vicinity of the
cilia. Thus, the bidirectional traction induced by motile cilia
inherently leads to a partial compartmentalization of the ventricular
system.
